Numerical simulations of the Cordilleran ice sheet through the last glacial cycle

### 2015-11-11 reviews/response-jarosch.tex

\documentclass[10pt]{article}

\usepackage[T1]{fontenc}
\usepackage[pdftex]{graphicx}
\usepackage[pdftex]{color}

\sloppy
\setlength{\parskip}{1.2ex}
\setlength{\parindent}{0em}
\renewcommand\familydefault{\sfdefault}

\graphicspath{{../figures/}}
\definecolor{journalname}{rgb}{0.34,0.59,0.82}

\begin{document}
\textbf{Authors' response to A.~H.~Jarosch}
\bigskip

% ----------------------------------------------------------------------
% Interactive comment text begins
% ----------------------------------------------------------------------

\newcommand{\sechead}[1]{\bigskip\noindent\textbf{#1}}
\newcommand{\referee}[1]{\bigskip\textcolor{journalname}{\textit{#1}}}
\newcommand{\msquote}[1]{\begin{quote}\textit{#1}\end{quote}}
\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}

To A.~H.~Jarosch,

% ----------------------------------------------------------------------

\sechead{1 \quad General comments}

\referee{%
    Seguinot et al. present in this well written and structured manuscript a
    numerical modelling study of the Cordilleran ice sheet through the last
    glacial cycle. The model is driven by several temperature reconstructions
    based on proxy data and model output is subsequently compared in detail to
    the existing geological evidence in the region. The study is of significant
    relevance as it focuses on the Cordilleran ice sheet evolution in the past,
    which is still poorly understood.}

\referee{%
    Nevertheless, the manuscript is quite unbalanced in its presentation as it
    focuses strongly on section~4 (Comparison with geological record) and by
    doing so neglects crucial details in section~2 (Model setup). This poses a
    fundamental challenge for understanding the science presented. If it is not
    quite clear what the model does and how it performs to start with, it
    becomes difficult to discuss the results of the modelling study and why
    there are mismatches with geological evidence.}

\referee{%
    An overall sensitivity study of the parameters used in the model is
    completely lacking, thereby making it almost impossible to understand
    different responses of the ice sheet model. After reading the manuscript,
    one is left with the impression that the authors assume the PISM ice sheet
    model to be a black box which just requires one initial ``correct'' setup
    with literature values. This notion is reflected in the current manuscript,
    where almost all mismatches of model output with geological evidence (as
    discussed in section~4) are attributed to climate variations lacking in the
    proxy data, or climate-ice sheet feedback mechanisms not represented in the
    model chain. Similarly in a previous study Seguinot et al. ( 2014 ) have
    focused only on the driving climate sensitivities and have omitted
    influences of the ice sheet model as well as mass balance model pa-
    rameters even though they note in that study that these sensitivities
    require attention as well.}

\referee{%
    What I advocate at this point is not a complete, strict sensitivity study
    of all parameters involved in the model setup (that would be probably a
    work package large enough to fill a science career). However several key
    parameters can be investigated with not too much effort. Contrasting the
    influence of e.g. basal sliding and ice rheology parameters with the
    influence of driving climate on the model results would help to estimate
    the overall sensitivity of the model system as well as help guiding future
    efforts performing such modelling studies. Implicitly the authors assume
    that all other model sensitivities are negligibly small in comparison to
    the driving climate. However it is obvious from an ice sheet model
    perspective that at least chosen basal sliding parameters as well as ice
    rheology parameters will strongly influence the shape and volume of the
    modelled ice sheet. Thus it would be nice to see evidence supporting the
    claim that driving climate is the only input to worry about being presented
    in the current manuscript. Or should it turn out that basal sliding and ice
    rheology play an important role too, as one would expect, then the relative
    importance of each including error estimates on the chosen parameters
    should be presented as well.}

\referee{%
    Generally section~4 appears to be quite long and seems to re-summarize
    known geological evidence for the region. At times the language is quite
    speculative, for example P4162 L1 and 7, P4164 L18, P4171 L7, L9, L17, L19
    and L20 and so forth. I would recommend to shorten that section to focus
    only on the geological evidence which can or can not clearly be reproduced
    by the presented model and avoid extensive speculation on what the reasons
    for mismatch are, especially in the present form of the manuscript, where a
    sensitivity study of the model itself is completely missing. However I
    leave the choice of how much geological evidence is discussed in the
    manuscript entirely up to the authors.}

% ----------------------------------------------------------------------

\sechead{2 \quad Specific Comments}

\referee{%
    I refer to text locations in the discussion paper by page number (P) and
    line number range (L) for the specific comments.}

\referee{%
    \textbf{P4151 L11-16:}
    In this sentence the authors refer back to their previous work ( Seguinot
    et al. , 2014 ) and highlight that the NARR temperature and precipitation
    fields are the most suitable present day climate datasets to be used.
    Especially since the NARR precipitation fields include steep precipitation
    gradients which are required as identified by Seguinot et al. ( 2014 ).
    NARR is delivered on a 32\,km Lambert grid, and thus it is ques- tionable
    how ``steep'' these gradients can be, given the rather smooth
    representation of the existing topography on a 32\,km grid. Seguinot et al.
    ( 2014 ) have partly discussed that however. NARR precipitation and
    temperature fields have been evaluated in detail based on available station
    data for large parts of the study domain dealt with in this manuscript.
    This evaluation ( Jarosch et al. , 2012 ) demonstrated that NARR has
    difficulties simulating orographic processes in the Coast Mountains which
    in turn results in unrealistic atmospheric conditions over the Rocky
    Mountains. Jarosch et al. ( 2012 ) further concluded that physics based
    downscaling is required to adequately drive glacier models in that region.
    The authors should argue in more detail here why they think that NARR
    precipitation fields at 32\,km are adequate to drive their model and
    reflect their arguments with the findings of Jarosch et al. ( 2012 ). A
    solid argument here is of special importance as the authors assume the
    present day precipitation fields to be valid throughout their model time
    period (120\,ky to present) without further corrections (cf. section~2.4
    equation~6).}

\referee{%
    \textbf{P4152 L11:}
    Basal topography is ``derived'' from ETOPO1 data. What does this mean? Do
    the authors just re-sample the DEM data to their 10\,km and 5\,km model
    grids (P4152 L21-22) or is there more processing done? The ETOPO1 data
    contains the present day ice volumes within the study region. Clarke et al.
    ( 2013 ) have estimated the ice volume in parts of that region to be $2530
    \pm 220\,km^2$, with maximum ice thicknesses up to 200\,m. It can be argued
    that the volume is negligible in this study and the authors should do so if
    they think this is appropriate, but I wonder about the ice thicknesses.
    Assuming that the authors did not remove the present day ice cover, basal
    topography could be up to 200\,m higher that it actually is in reality.
    Given their used temperature lapse rate of $6\,K\,km^{-1}$ (P4157 L1),
    parts of the Cordilleran ice sheet growing in those regions with 200\,m too
    high topography would experience a 1.2\,K colder atmosphere than it
    actually should in reality. This favours unrealistic ice growth and thus
    the omission of present day ice cover removal should be clearly argued for
    in the manuscript.}

\referee{%
    \textbf{P4153 L2-3:}
    That the ``shallow shelf approximation'' (SSA) is used as a ``sliding law''
    for the ``shallow ice approximation'' (SIA) is a confusing statement in
    this context. Bueler has coined the term in his 2009 paper as cited in the
    manuscript. However the casual reader will be confused at this point,
    especially since the authors state the pseudoplastic sliding law the model
    actually uses in equation~1. I would recommend to leave out the statement
    on the SSA being the ``sliding law'' for the SIA.}

\referee{%
    \textbf{P4153 L5-6:}
    As stated here, ice rheology within the used ice sheet model is based on
    Aschwanden et al. ( 2012 ). This enthalpy based formulation has proofed
    itself to be very suitable for estimating ice rheology in ice sheet models,
    but it also depends on several parameters to translate enthalpy within the
    ice to ice viscosity (see Aschwanden et al. ( 2012 ), equations 62-65). The
    authors do not mention any of these parameters (e.g. any of the rate
    factors or nonlinear power $n$ ) within the manuscript or in Table~1. I
    have mentioned above in the general comments section that parameters used
    in ice rheology and basal sliding formulations are important model
    parameters which will influence the ice sheet model output and that a basic
    sensitivity study on those parameters is required to understand the model
    results. Here the authors could start with listing the parameters used in
    the ice rheology formulation, than continue with estimating uncertainties
    for those from literature and afterwards perform additional model
    simulations to identify the influence of the chosen parameter sets on the
    ice volume and ice margin position history the model creates. In the end
    the authors will be able to identify the relative importance of
    uncertainties in driving climate as well as model parameters, which will
    strengthen their discussion in section~4.}

\referee{%
    \textbf{P4153 L8:}
    It is not clear where the geothermal heat flux boundary is located. Does
    the ``depth of 3\,km'' refer to a depth measured from the ice surface,
    which would not make much sense for a ice thickness evolving ice sheet
    model or is it measured from the ice-bedrock interface downward. In that
    case the term ``computed subglacially'' is confusing as it refers to the
    ice-bedrock interface. Please be more specific here.}

\referee{%
    \textbf{P4153 L16 - P4154 L0:}
    Here the authors describe the basal sliding setup in their model. However
    they do not explain how they came up with the parameters used in
    equations~1-3 that are listed in Table~1 (part on ``basal sliding''). What
    motivates these parameter choices (references?) and how sensitive is the
    model and its results to these choices? Both question come instantly to
    mind and need to be addressed in detail. Here a basic sensitivity study on
    how basal sliding parameters in the model control the outcome discussed in
    section~4 is in order and I strongly recommend to include one in the
    manuscript. The authors can start by estimating the uncertainties in the
    chosen basal sliding parameters and run two extra simulation runs with
    their preferred climate forcing and the end member values of the
    uncertainties. This would create the most simple sensitivity study with
    respect to basal sliding, but would be extremely helpful for the argument
    made above in my general comments.}

\referee{%
    \textbf{P4156 L3-5:}
    In addition to what I have stated above on the NARR precipitation fields
    and their suitability, it is important to state at this location in the
    manuscript how the 32\,km NARR data is translated to the 10 and 5\,km
    computational grids of the current study. I disagree with the notion that a
    32\,km precipitation field can be called ``high- resolution'' in the
    context of 10 and 5\,km grid based ice sheet modelling. The input data is
    either 3 or 6 times coarser than the numerical grid, thus not at all
    high-resolution. Seguinot et al. ( 2014 ) state in their section~3.3 that
    the NARR data fields have been bilinearly interpolated to 10\,km resolution
    in their work. Did the authors do the same here for their 10 and 5\,km
    working grids? This is crucial information to be included in the
    manuscript. It has been demonstrated by spectral power analysis ( Jarosch
    et al. , 2012 ) that the NARR precipitation fields do not contain any
    significant spacial information below approximately 39\,km resolution and
    that bilinear interpolation does not add any information whatsoever on
    smaller scales, which should come to no surprise. Physics-based downscaling
    techniques however are able to add spatial information to precipitation
    fields down to about 1\,km grid sizes ( Jarosch et al. , 2012 ). Taking
    these findings into the current context of the manuscript at hand, the NARR
    precipitation fields can hardly be called ``high resolution'' with their
    effective precipitation grid size of 39\,km. The authors should argue for
    their choice of not performing any downscaling whatsoever to their
    computational grids of 10 and 5\,km for precipitation and temperature and
    discuss their choice in the light of the findings from Jarosch et al. (
    2012 ). Temperature however is better constrained in NARR ( Jarosch et al.
    , 2012 ) and contains spectral information down to 10\,km resolution, which
    justifies the usage of NARR temperature fields on the 10\,km computational
    grid of this study. The 5\,km grid still needs to be argued for.}

\referee{%
    \textbf{P4157 L1:}
    How is a fixed temperature lapse rate justified for simulations over 120k
    years, when there is ample published evidence that temperature lapse rates
    vary significantly within space and time? I am sure that the choice of
    $\gamma$ in this study has a significant influence on the model outcome and
    I leave it to the authors to explore this possibility.}

% ----------------------------------------------------------------------

\sechead{3 \quad Technical Corrections}

\referee{%
    \textbf{P4161 L10:}
    ``further analysis further;'' maybe change to ``further analysis'' or
    ``further analysis here''.}

\referee{%
    \textbf{P4166 L13:}
    double ``the'' in the sentence.}

\referee{%
    I hope the authors find my comments helpful in revising their manuscript
    and wish them success for their future endeavours.}

% ----------------------------------------------------------------------
% Interactive comment text ends
% ----------------------------------------------------------------------

\end{document}

### 2015-12-16 cordillera-cycle.tex
Restore manuscript format.
Rename all figures. Fix Makefile.

% Copernicus manuscript
%\documentclass[tc, manuscript]{copernicus}

% Copernicus final print
\documentclass[tc]{copernicus}

@@

@@

\begin{table*}[t]
  \caption{Palaeo-temperature proxy records and scaling parameters yielding
           temperature offset time-series used to force the ice sheet model
           through the last glacial cycle (Fig.~\ref{fig:timeseries}). $f$
           corresponds to the scaling factor adopted to yield last glacial
           maximum ice limits in the vicinity of mapped end moraines, and
           $T_{[32, 22]}$ refers to the resulting mean temperature anomaly
           during the period -32 to~-22~ka after scaling.}
  \label{tab:records}
  {\begin{tabular}{l|ccc|ccc|l}
    \tophline
    Record & Latitude & Longitude & Elev. & Proxy & $f$ & $T_{[32, 22]}$
           & Reference\\
    & & & (m~a.s.l.) & & & (K) & \\
    \middlehline
    GRIP     &  72{\degree} 35' N  % 72.58 (decimal)
             &  37{\degree} 38' W  % 37.64 (decimal)
             & 3238
             & \chem{\delta^{18}O}
             & 0.37 & -5.8  % -16.4126 (before scaling)
             & \citet{Dansgaard.etal.1993} \\

@@

Temperature offset time-series, ${\Delta}T_{TS}$, are derived from
palaeo-temperature proxy records from
the Greenland Ice Core Project \citep[GRIP,][]{Dansgaard.etal.1993}, the
North Greenland Ice Core Project \citep[NGRIP,][]{Andersen.etal.2004},
the European Project for Ice Coring in Antarctica \citep[EPICA,][]
{Jouzel.etal.2007}, the Vostok ice core \citep{Petit.etal.1999}, and Ocean
Drilling Program (ODP) sites 1012 and 1020, both located off the coast of
California \citep{Herbert.etal.2001}. Palaeo-temperature anomalies from the
GRIP and NGRIP
records were calculated from oxygen isotope (\chem{\delta^{18}O}) measurements
using a quadratic equation \citep{Johnsen.etal.1995},
\begin{align}
    {\Delta}T_{TS}(t) = &-11.88 [\chem{\delta^{18}O}(t)
                                -\chem{\delta^{18}O}(0)] \nonumber \\
                        &-0.1925[\chem{\delta^{18}O}(t)^2
                                 -\chem{\delta^{18}O}(0)^2] \,,
\end{align}
while temperature reconstructions from Antarctic and oceanic cores were
provided as such. All records were scaled linearly (Table~\ref{tab:records}) in
order to simulate comparable ice extents at the LGM  (Table~\ref{tab:extrema})
and realistic outlines (Fig.~\ref{fig:snapshots}).

@@

\begin{figure*}
  \includegraphics{timeseries}
  \caption{Temperature offset time-series from ice core and ocean records
           (Table~\ref{tab:records}) used as palaeo-climate forcing for the ice
           sheet model (top panel), and modelled ice volume (bottom panel)
           through the last 120\,ka. Ice volumes are expressed in meters of sea
           level equivalent (m~s.l.e.). Gray fields indicate Marine Oxygen
           Isotope Stage (MIS) boundaries for MIS~2 and MIS~4 according to a
           global compilation of benthic \chem{\delta^{18}O} records
           \citep{Lisiecki.Raymo.2005}. Hatched rectangles highlight the
           time-volume span for ice volume extremes corresponding to MIS~4
           (61.9--56.5\,ka), MIS~3 (53.0--41.3\,ka), and MIS~2 (LGM,
           23.2--16.8\,ka). Dotted lines correspond to GRIP- and EPICA-driven
           5\,km-resolution runs.}
  \label{fig:timeseries}
\end{figure*}

\begin{figure*}
  \includegraphics{snapshots}
  \caption{Snapshots of modelled surface topography (500\,m contours)
           corresponding to the ice volume extremes indicated on
           Fig.~\ref{fig:timeseries}. An ice cap persists over the Skeena
           Mountains (SM) during MIS~3. Note the occurence of spatial
           similarities despite large differences in timing.}
  \label{fig:snapshots}
\end{figure*}

\begin{table*}
  \caption{Extremes in Cordilleran ice sheet volume and extent corresponding to
           MIS~4, 3 and 2 for each of the six low-resolution simulations
           (Fig.~\ref{fig:timeseries}).}
  \label{tab:extrema}
  {\begin{tabular}{l*{3}{|ccc}}
    \tophline
             & \multicolumn{3}{c}{Age (ka)}
             & \multicolumn{3}{c}{Ice extent (\unit{10^6\,km^2})}
             & \multicolumn{3}{c}{Ice volume (m~s.l.e.)} \\
    Record   &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2 \\
    \middlehline
    GRIP     &  57.45 &  42.93 &  19.14
             &   1.87 &   0.69 &   2.07
             &   7.35 &   1.69 &   8.71 \\
    NGRIP    &  60.25 &  45.86 &  22.85
             &   2.11 &   0.74 &   2.10
             &   8.84 &   1.82 &   8.76 \\
    EPICA    &  61.89 &  45.30 &  17.23
             &   1.51 &   1.01 &   2.09
             &   5.11 &   2.74 &   8.91 \\
    Vostok   &  60.86 &  41.25 &  16.84
             &   1.51 &   1.03 &   2.01
             &   5.15 &   2.88 &   8.40 \\
    ODP 1012 &  56.46 &  47.39 &  23.19
             &   1.38 &   0.90 &   2.07
             &   4.39 &   2.35 &   8.62 \\
    ODP 1020 &  60.23 &  52.97 &  20.56
             &   1.25 &   0.72 &   2.09
             &   3.84 &   1.75 &   8.81 \\
    \middlehline
    Minimum  &  61.89 &  52.97 &  23.19
             &   1.25 &   0.69 &   2.01
             &   3.84 &   1.69 &   8.40 \\
    Maximum  &  56.46 &  41.25 &  16.84
             &   2.11 &   1.03 &   2.10
             &   8.84 &   2.88 &   8.91 \\
    \bottomhline
  \end{tabular}}
  \belowtable{}
\end{table*}

Despite large differences in the input climate forcing
(Fig.~\ref{fig:timeseries}, upper panel), model output presents consistent
features that can be observed across the range of forcing data used. In all
simulations, modelled ice volumes remain relatively low during most of the
glacial cycle, except during two major glacial events which occur between 61.9
and 56.5\,ka during MIS~4, and between 23.2 and 16.8\,ka during MIS~2
(Fig.~\ref{fig:timeseries}, lower panel). An ice volume minimum is
consistently reached between 53.0 and 41.3\,ka during MIS~3. However, the
magnitude and precise timing of these three events depend significantly on the
choice of proxy record used to derive a time-dependent climate forcing
(Table~\ref{tab:extrema}).

@@

While simulations driven by the GRIP and the two Antarctic palaeo-temperature
records attain a last ice volume maximum between 19.1 and 16.8\,ka, those
informed by the NGRIP and the two oceanic palaeo-temperature records attain their
maximum ice volumes thousands of years earlier. Moreover, the ODP~1012 run
yields a rapid deglaciation of the modelled area prior to 17\,ka. The ODP~1020
simulation predicts an early maximum in ice volume at 23.2\,ka, followed by
slower deglaciation than modelled using the other palaeo-temperature records.
Finally, whereas model runs forced by the Antarctic palaeo-temperature records
result in a rapid and uninterrupted deglaciation after the LGM, the
simulation driven by the GRIP palaeo-temperature record also results in a rapid
deglaciation but in three steps, separated by two periods of ice sheet regrowth
(Fig.~\ref{fig:timeseries}).

@@

Despite large differences in the timing of attained volume extrema
(Table~\ref{tab:extrema}), all model runs show relatively consistent
patterns of glaciation. During MIS~4, all simulations produce an extensive ice
sheet, covering an area of at least half of that attained during MIS~2
(Table~\ref{tab:extrema}; Fig.~\ref{fig:snapshots}, upper panels).
Corresponding maximum ice volumes also differ significantly between model runs,
and vary between 3.84 and~8.84\,m sea level equivalents (m.~s.l.e.;
Table~\ref{tab:extrema}).

In the MIS~3 ice volume minimum reconstructions, a central ice cap persists
over the Skeena
Mountains (Fig.~\ref{fig:snapshots}, middle panels). Although this ice cap is
present in all simulations, its dimensions depend sensitively on the choice of
the applied palaeo-temperature record. Modelled ice volume minima spread over a
wide range between 1.69 and 2.88\,m~s.l.e. (Table~\ref{tab:extrema}).

Modelled ice sheet geometries during the LGM (MIS~2; Fig.~\ref{fig:snapshots},
lower panels) invariably include a ca. 1500\,km-long central divide above
3000\,m~a.s.l. located along the spine of the Rocky Mountains. Although the
similarity of modelled ice extents is a direct result of the choice of scaling
factors applied to different palaeo-temperature proxy records
(Table~\ref{tab:records}), it is interesting to note that modelled maximum ice
volumes also fall within a tight range of 8.40 to 8.91\,m~s.l.e.
(Table~\ref{tab:extrema}).

@@

Independently of the palaeo-temperature records
used to force the ice sheet model, our simulations consistently produce two
glacial maxima during the last glacial cycle. The first maximum configuration
is obtained during MIS~4 (61.9--56.5\,ka) and the second during MIS~2
(23.2--16.8\,ka; Figs.~\ref{fig:timeseries}, \ref{fig:snapshots};
Table~\ref{tab:extrema}). These events broadly correspond in timing to the
Gladstone
(MIS~4) and McConnell (MIS~2) glaciations documented by geological evidence for
the northern sector of the Cordilleran ice sheet
    \citep{Duk-Rodkin.etal.1996, Ward.etal.2007,
           Stroeven.etal.2010, Stroeven.etal.2014},
and to the Fraser Glaciation (MIS~2) documented for its southern sector
    \citep{Porter.Swanson.1998, Margold.etal.2014}.
There is stratigraphical evidence for an MIS~4 glaciation in British Columbia
\citep{Clague.Ward.2011} and in the Puget Lowland \citep{Troost.2014}, but
their extent and timing are still highly conjectural
    \citep[perhaps MIS~4 or early MIS~3; e.g.,][]{Cosma.etal.2008}.

@@

Among the simulations presented here, only those forced with the GRIP, EPICA
and Vostok palaeo-temperature records yield Cordilleran ice sheet maximum
extents that may be compatible with these field constraints
(Fig.~\ref{fig:timeseries}, lower panel; Table~\ref{tab:extrema}).
Simulations driven by the NGRIP, ODP~1012 and ODP~1020
palaeo-temperature records, on the contrary, yield MIS~2 maximum Cordilleran
ice sheet volumes that pre-date field-based constraints by several thousands of
years (about 6, 6 and 4\,ka respectively).
Concerning the simulations driven by oceanic records, this early
deglaciation is caused by an early warming present in the alkenone
palaeo-temperature reconstructions (Fig.~\ref{fig:timeseries}, upper panel;
\citealp[Fig.~3]{Herbert.etal.2001}). However, this
early warming is a local effect, corresponding to a weakening of the California
current \citep{Herbert.etal.2001}. The California current, driving cold
waters southwards along the south-western coast of North America, has been
shown to have weakened during each peak of global glaciation (in SPECMAP)
during the past 550\,ka, including the LGM, resulting in paradoxically warmer
sea-surface temperatures at the locations of the ODP~1012 and ODP~1020 sites
\citep{Herbert.etal.2001}.

>>>>

% Copernicus manuscript
\documentclass[tc, manuscript]{copernicus}

% Copernicus final print
%\documentclass[tc]{copernicus}

@@

% Push floats to the end
\usepackage[nolists, nomarkers]{endfloat}

@@

\begin{table*}[t]
  \caption{Palaeo-temperature proxy records and scaling parameters yielding
           temperature offset time-series used to force the ice sheet model
           through the last glacial cycle (Fig.~\ref{fig:lr_ts}). $f$
           corresponds to the scaling factor adopted to yield last glacial
           maximum ice limits in the vicinity of mapped end moraines, and
           $T_{[32, 22]}$ refers to the resulting mean temperature anomaly
           during the period -32 to~-22~ka after scaling.}
  \label{tab:records}
  {\begin{tabular}{l|ccc|ccc|l}
    \tophline
    Record & Latitude & Longitude & Elev. & Proxy & $f$ & $T_{[32, 22]}$
           & Reference\\
    & & & (m~a.s.l.) & & & (K) & \\
    \middlehline
    GRIP     &  72{\degree} 35' N  % 72.58 (decimal)
             &  37{\degree} 38' W  % 37.64 (decimal)
             & 3238
             & \chem{\delta^{18}O}
             & 0.37 & -5.8  % -16.4126 (before scaling)
             & \citet{Dansgaard.etal.1993} \\

@@

Temperature offset time-series, ${\Delta}T_{TS}$, are derived from
palaeo-temperature proxy records from
the Greenland Ice Core Project \citep[GRIP,][]{Dansgaard.etal.1993}, the
North Greenland Ice Core Project \citep[NGRIP,][]{Andersen.etal.2004},
the European Project for Ice Coring in Antarctica \citep[EPICA,][]
{Jouzel.etal.2007}, the Vostok ice core \citep{Petit.etal.1999}, and Ocean
Drilling Program (ODP) sites 1012 and 1020, both located off the coast of
California \citep{Herbert.etal.2001}. Palaeo-temperature anomalies from the
GRIP and NGRIP
records were calculated from oxygen isotope (\chem{\delta^{18}O}) measurements
using a quadratic equation \citep{Johnsen.etal.1995},
\begin{align}
    {\Delta}T_{TS}(t) = &-11.88 [\chem{\delta^{18}O}(t)
                                -\chem{\delta^{18}O}(0)] \nonumber \\
                        &-0.1925[\chem{\delta^{18}O}(t)^2
                                 -\chem{\delta^{18}O}(0)^2] \,,
\end{align}
while temperature reconstructions from Antarctic and oceanic cores were
provided as such. All records were scaled linearly (Table~\ref{tab:records}) in
order to simulate comparable ice extents at the LGM  (Table~\ref{tab:extrema})
and realistic outlines (Fig.~\ref{fig:lr_maps}).

@@

\begin{figure*}
  \includegraphics{lr_ts}
  \caption{Temperature offset time-series from ice core and ocean records
           (Table~\ref{tab:records}) used as palaeo-climate forcing for the ice
           sheet model (top panel), and modelled ice volume (bottom panel)
           through the last 120\,ka. Ice volumes are expressed in meters of sea
           level equivalent (m~s.l.e.). Gray fields indicate Marine Oxygen
           Isotope Stage (MIS) boundaries for MIS~2 and MIS~4 according to a
           global compilation of benthic \chem{\delta^{18}O} records
           \citep{Lisiecki.Raymo.2005}. Hatched rectangles highlight the
           time-volume span for ice volume extremes corresponding to MIS~4
           (61.9--56.5\,ka), MIS~3 (53.0--41.3\,ka), and MIS~2 (LGM,
           23.2--16.8\,ka). Dotted lines correspond to GRIP- and EPICA-driven
           5\,km-resolution runs.}
  \label{fig:lr_ts}
\end{figure*}

\begin{figure*}
  \includegraphics{lr_maps}
  \caption{Snapshots of modelled surface topography (500\,m contours)
           corresponding to the ice volume extremes indicated on
           Fig.~\ref{fig:lr_ts}. An ice cap persists over the Skeena
           Mountains (SM) during MIS~3. Note the occurence of spatial
           similarities despite large differences in timing.}
  \label{fig:lr_maps}
\end{figure*}

\begin{table*}
  \caption{Extremes in Cordilleran ice sheet volume and extent corresponding to
           MIS~4, 3 and 2 for each of the six low-resolution simulations
           (Fig.~\ref{fig:lr_ts}).}
  \label{tab:extrema}
  {\begin{tabular}{l*{3}{|ccc}}
    \tophline
             & \multicolumn{3}{c}{Age (ka)}
             & \multicolumn{3}{c}{Ice extent (\unit{10^6\,km^2})}
             & \multicolumn{3}{c}{Ice volume (m~s.l.e.)} \\
    Record   &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2 \\
    \middlehline
    GRIP     &  57.45 &  42.93 &  19.14
             &   1.87 &   0.69 &   2.07
             &   7.35 &   1.69 &   8.71 \\
    NGRIP    &  60.25 &  45.86 &  22.85
             &   2.11 &   0.74 &   2.10
             &   8.84 &   1.82 &   8.76 \\
    EPICA    &  61.89 &  45.30 &  17.23
             &   1.51 &   1.01 &   2.09
             &   5.11 &   2.74 &   8.91 \\
    Vostok   &  60.86 &  41.25 &  16.84
             &   1.51 &   1.03 &   2.01
             &   5.15 &   2.88 &   8.40 \\
    ODP 1012 &  56.46 &  47.39 &  23.19
             &   1.38 &   0.90 &   2.07
             &   4.39 &   2.35 &   8.62 \\
    ODP 1020 &  60.23 &  52.97 &  20.56
             &   1.25 &   0.72 &   2.09
             &   3.84 &   1.75 &   8.81 \\
    \middlehline
    Minimum  &  61.89 &  52.97 &  23.19
             &   1.25 &   0.69 &   2.01
             &   3.84 &   1.69 &   8.40 \\
    Maximum  &  56.46 &  41.25 &  16.84
             &   2.11 &   1.03 &   2.10
             &   8.84 &   2.88 &   8.91 \\
    \bottomhline
  \end{tabular}}
  \belowtable{}
\end{table*}

Despite large differences in the input climate forcing
(Fig.~\ref{fig:lr_ts}, upper panel), model output presents consistent
features that can be observed across the range of forcing data used. In all
simulations, modelled ice volumes remain relatively low during most of the
glacial cycle, except during two major glacial events which occur between 61.9
and 56.5\,ka during MIS~4, and between 23.2 and 16.8\,ka during MIS~2
(Fig.~\ref{fig:lr_ts}, lower panel). An ice volume minimum is
consistently reached between 53.0 and 41.3\,ka during MIS~3. However, the
magnitude and precise timing of these three events depend significantly on the
choice of proxy record used to derive a time-dependent climate forcing
(Table~\ref{tab:extrema}).

@@

While simulations driven by the GRIP and the two Antarctic palaeo-temperature
records attain a last ice volume maximum between 19.1 and 16.8\,ka, those
informed by the NGRIP and the two oceanic palaeo-temperature records attain their
maximum ice volumes thousands of years earlier. Moreover, the ODP~1012 run
yields a rapid deglaciation of the modelled area prior to 17\,ka. The ODP~1020
simulation predicts an early maximum in ice volume at 23.2\,ka, followed by
slower deglaciation than modelled using the other palaeo-temperature records.
Finally, whereas model runs forced by the Antarctic palaeo-temperature records
result in a rapid and uninterrupted deglaciation after the LGM, the
simulation driven by the GRIP palaeo-temperature record also results in a rapid
deglaciation but in three steps, separated by two periods of ice sheet regrowth
(Fig.~\ref{fig:lr_ts}).

@@

Despite large differences in the timing of attained volume extrema
(Table~\ref{tab:extrema}), all model runs show relatively consistent
patterns of glaciation. During MIS~4, all simulations produce an extensive ice
sheet, covering an area of at least half of that attained during MIS~2
(Table~\ref{tab:extrema}; Fig.~\ref{fig:lr_maps}, upper panels).
Corresponding maximum ice volumes also differ significantly between model runs,
and vary between 3.84 and~8.84\,m sea level equivalents (m.~s.l.e.;
Table~\ref{tab:extrema}).

In the MIS~3 ice volume minimum reconstructions, a central ice cap persists
over the Skeena
Mountains (Fig.~\ref{fig:lr_maps}, middle panels). Although this ice cap is
present in all simulations, its dimensions depend sensitively on the choice of
the applied palaeo-temperature record. Modelled ice volume minima spread over a
wide range between 1.69 and 2.88\,m~s.l.e. (Table~\ref{tab:extrema}).

Modelled ice sheet geometries during the LGM (MIS~2; Fig.~\ref{fig:lr_maps},
lower panels) invariably include a ca. 1500\,km-long central divide above
3000\,m~a.s.l. located along the spine of the Rocky Mountains. Although the
similarity of modelled ice extents is a direct result of the choice of scaling
factors applied to different palaeo-temperature proxy records
(Table~\ref{tab:records}), it is interesting to note that modelled maximum ice
volumes also fall within a tight range of 8.40 to 8.91\,m~s.l.e.
(Table~\ref{tab:extrema}).

@@

Independently of the palaeo-temperature records
used to force the ice sheet model, our simulations consistently produce two
glacial maxima during the last glacial cycle. The first maximum configuration
is obtained during MIS~4 (61.9--56.5\,ka) and the second during MIS~2
(23.2--16.8\,ka; Figs.~\ref{fig:lr_ts}, \ref{fig:lr_maps};
Table~\ref{tab:extrema}). These events broadly correspond in timing to the
Gladstone
(MIS~4) and McConnell (MIS~2) glaciations documented by geological evidence for
the northern sector of the Cordilleran ice sheet
    \citep{Duk-Rodkin.etal.1996, Ward.etal.2007,
           Stroeven.etal.2010, Stroeven.etal.2014},
and to the Fraser Glaciation (MIS~2) documented for its southern sector
    \citep{Porter.Swanson.1998, Margold.etal.2014}.
There is stratigraphical evidence for an MIS~4 glaciation in British Columbia
\citep{Clague.Ward.2011} and in the Puget Lowland \citep{Troost.2014}, but
their extent and timing are still highly conjectural
    \citep[perhaps MIS~4 or early MIS~3; e.g.,][]{Cosma.etal.2008}.

@@

Among the simulations presented here, only those forced with the GRIP, EPICA
and Vostok palaeo-temperature records yield Cordilleran ice sheet maximum
extents that may be compatible with these field constraints
(Fig.~\ref{fig:lr_ts}, lower panel; Table~\ref{tab:extrema}).
Simulations driven by the NGRIP, ODP~1012 and ODP~1020
palaeo-temperature records, on the contrary, yield MIS~2 maximum Cordilleran
ice sheet volumes that pre-date field-based constraints by several thousands of
years (about 6, 6 and 4\,ka respectively).
Concerning the simulations driven by oceanic records, this early
deglaciation is caused by an early warming present in the alkenone
palaeo-temperature reconstructions (Fig.~\ref{fig:lr_ts}, upper panel;
\citealp[Fig.~3]{Herbert.etal.2001}). However, this
early warming is a local effect, corresponding to a weakening of the California
current \citep{Herbert.etal.2001}. The California current, driving cold
waters southwards along the south-western coast of North America, has been
shown to have weakened during each peak of global glaciation (in SPECMAP)
during the past 550\,ka, including the LGM, resulting in paradoxically warmer
sea-surface temperatures at the locations of the ODP~1012 and ODP~1020 sites
\citep{Herbert.etal.2001}.
